---
title: "Visualization"
output: github_document
date: "2026-10-01"
---

```{r setup}
library(tidyverse)
library(ggridges)

library(p8105.datasets)
data("weather_df")
```

```{r}
weather_df #365 days, 2 years, 3 weather stations
```

Let's make a scatterplot!
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()

# the first thing in the argument is the name of the dataset
# x = the variable you want on the x axis
# y = the variable you want on the y axis
# this creates a scatter plot
```

He likes the dataframe first.
```{r}
ggp_temp_scatterplot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_temp_scatterplot

# this does the same thing as the first chunk
```

Let's make this a bit fancier...
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .25) + #alpha makes points a bit transparent
  geom_smooth(se = FALSE) #adds a smooth line through the points; gives error bars by default so se=F gets rid of them 

# "color = name" makes the data points that correspond to the different values of the variable "name" different colors, and creates a legend to show what values the colors correspond to 
```

```{r}
#only 1 smooth line
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .25) + #alpha makes points a bit transparent; adding aes makes it so that colors are only defined for points and won't show up in geom_smooth
  geom_smooth(se = FALSE) #adds a smooth line through the points; gives error bars by default so se=F gets rid of them 
```
the aesthetics are up to you

```{r}
#no points
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE) 
```

Show faceting
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  facet_grid(. ~ name) #nothing separating rows, name separates cols; switching to facet_grid(name ~ .) leads to graphs on top of each other
```

```{r}
#does same thing as above, just a diff way to have column specification
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  facet_grid(cols = vars(name))
```

Let's look at something else.

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

Make a plot of Central Park tmax and tmin only, and convert temperatures o farenheit

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax = tmax * (9/5) + 32,
    tmin = tmin * (9/5) + 32
  ) |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

what's a hex plot

```{r}
weather_df |> 
  ggplot(aes(x=tmin, y=tmax))+
  geom_hex() #gives an idea of the data summary and what it looks like. more shading = more common
```

## Univariate plots

```{r}
weather_df |> 
  ggplot(aes(x = tmax)) + 
  geom_histogram()
```


```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill=name)) + #color = boundary of bar instead of fill, so use fill instead
  geom_histogram(position = "dodge") # dodge makes it so they don't overlap
```
```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill=name)) + #color = boundary of bar instead of fill, so use fill instead
  geom_histogram()+
  facet_grid(. ~ name) # better than dodge
```

Density plots are great!!

```{r}
weather_df |> 
  ggplot(aes(x=tmax, color=name)) +
  geom_density()

weather_df |> 
  ggplot(aes(x=tmax, fill=name)) +
  geom_density(alpha= .3)
```
Boxplots
```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) + #adding x separates out the 3 boxplots by name
  geom_boxplot()
```

Violin plots
```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin() # shows distribution info (density plots on the side and mirrored); like a combined density and box plots
```

Ridge plots
```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges() #separates curves horizontally; helpful if you have a lot of categories
```

## Save some of the plots

```{r}
ggp_weather = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.5) +
  facet_grid(.~name)

ggp_weather

ggsave("images/ggp_weather.pdf", ggp_weather)
```

```{r fig.width = 12, fig.height = 12}
weather_df |> 
  ggplot(aes(x=tmin, y = tmax))+
  geom_point()
```

to fix the size of the images, can use: 

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)